### 2015-12-04 body/equilibrium.tex
corrections andrea

\section{Heuristics for relevant contact selection}

\subsection{A heuristic for robust static equilibrium}
The planner is designed so that any contact configuration it outputs is in static equilibrium.
Equilibrium is of course critical in multipedal locomotion. For this reason we are interested in a robust 
criterion, that ensures that the robot remains balanced in a real world application, regardless of the errors
that can result from perception and control uncertainties.

We first give a linear program (LP) that asserts that a contact configuration is in static equilibrium.
From this formulation we derive a new LP that quantifies the robustness of the configuration regarding equilibrium.
From this value we can in turn either define a heuristic to choose the most robust candidate, or set a treshold on the required robustness
for any configuration. While the LP we propose is original, it is based on an analysis of the problem we proposed in \cite{Prete2016}, where the interested reader
can find more details.

\subsubsection{Conditions for static balance.}
We first define the variables of the problem, for a robot $R$ and $e$ contact points, expressed in world coordinates:
\begin{itemize}
\item $\mathbf{c} \in \mathbb{R}^3$ is the robot center of mass (COM);
\item $\mathbf{L}  \in \mathbb{R}^3 $ is the angular momentum at the COM;
\item $\mathbf{m} \in \mathbb{R}$ is the robot mass;
\item $\mathbf{g} = [0,0,-9.81]^T$ is the gravity acceleration;
\item $\mu$ is the friction coefficient;
\item for the i-th contact $1 \leq i \leq e$:
	\begin{itemize}
	\item $\mathbf{p_i}$ is the contact position;
	\item $\mathbf{f_i}$ is the force applied at $\mathbf{p_i}$;
	\item $\mathbf{n}_i,\mathbf{t}_{i1},\mathbf{t}_{i2}$ form a local cartesian coordinate system centered at $\mathbf{p_i}$. $\mathbf{n}_i$ is aligned
	with the contact surface normal, and the $\mathbf{t}_i$s are tangent vectors.
	\end{itemize}
\end{itemize}

According to Coulomb's law, the non slipping condition is verified if all the contact forces lie in the friction cone defined by the surface.
As classically done, we linearize the friction cone in a conservative fashion with a pyramid, described by four generating rays. We choose for instance:
\begin{equation*}
\mathbf{V}_{i} = \mat{\mathbf{n}_{i} + \mu \mathbf{t}_{i1} & \mathbf{n}_{i} -\mu \mathbf{t}_{i1} & \mathbf{n}_{i} + \mu \mathbf{t}_{i2} & \mathbf{n}_{i} - \mu \mathbf{t}_{i2}}^T
\end{equation*}

Any force belonging to the linearized cone
can thus be express as a convex combination of its four generating rays.
Therefore we can express the non slipping constraint on $\mathbf{f}_i$ as:

\begin{equation*}
\forall i  \qquad  \exists \bm{\beta}_i \in \mathbb{R}^{4} , \bm{\beta}_i \ge 0 \wedge \mathbf{f}_{i} = \mathbf{V}_{i} \bm{\beta}_i \qquad 
\end{equation*}

we can stack the constraints to obtain:
\begin{equation}\label{eq:gen}
\exists \bm{\beta} \in \mathbb{R}^{4e} ,  \bm{\beta} \ge 0 \wedge \mathbf{f} = \mathbf{V} \bm{\beta} \qquad 
\end{equation}
where $\mathbf{V} = \diag{ \{\mathbf{V}_1, \dots, \mathbf{V}_e\} }$, and $\mathbf{f} = [\mathbf{f}_0,...,\mathbf{f}_e]^T$.

From the Newton-Euler equations, and assuming a null acceleration, we also constrain the sum of contact forces to compensate the gravity at the center of mass:

%~ \begin{align}
%~ \underbrace{\mat{m (\ddot{\mathbf{c}} - \mathbf{g})  \\ m \mathbf{c} \times (\ddot{\mathbf{c}} - \mathbf{g}) + \dot{\mathbf{L}}}}_\mathbf{w}
%~ = 
%~ \underbrace{
%~ \mat{\mathbf{I}_3 & \dots & \mathbf{I}_3 \\
%~ \hat{\mathbf{p}}_1 & \dots & \hat{\mathbf{p}}_e} \mathbf{V}
%~ }_\mathbf{G}
%~ \bm{\beta},
%~ \end{align}
%~ where $\mathbf{w}\in \Rv{6}$ is the so-called gravito-inertial wrench (GIW) \citep{qiu:dhm:2011, Caron2015} and $\hat{\mathbf{p}} \in \R{3}{3}$ is the cross-product matrix associated to $\mathbf{p}$.

%~ Static balance assumes zero acceleration thus we can write:
\begin{align} \label{eq:new_eul}
\underbrace{
\mat{\mathbf{I}_3 & \dots & \mathbf{I}_3 \\
\hat{\mathbf{p}}_1 & \dots & \hat{\mathbf{p}}_e} \mathbf{V}
}_\mathbf{G} \bm{\beta}, = 
\underbrace{\mat{\mathbf{0}_{3\times 3} \\ m \hat{\mathbf{g}}}}_{\mathbf{D}} \mathbf{c} + 
\underbrace{\mat{-m\mathbf{g} \\ \mathbf{0}}}_{\mathbf{d}}
\end{align}
where $\hat{\mathbf{x}} \in \R{3}{3}$ is the cross-product matrix associated to $\mathbf{x}$.
The right side of \ref{eq:new_eul} is independent from $\mathbf{c}^z$ because of the cross product matrix $\hat{\mathbf{g}}$ inside $\mathbf{D}$, so we can rewrite it as:
\begin{align} \label{eq:new_eul2d}
\mathbf{w}_0 = \mathbf{D}^{xy} \mathbf{c}^{xy} + \mathbf{d}
\end{align}

If there exists a $\bm{\beta}^*$ satisfying equations \ref{eq:gen} and \ref{eq:new_eul2d}, it means that the configuration is statically balanced.
The problem can then be formulated as an LP:

\begin{equation} \label{eq:lin_prog} \begin{aligned}
\find \quad & \bm{\beta} \in \Rv{4e} \\
\st \quad &\mathbf{G} \bm{\beta} = \mathbf{D}^{xy} \mathbf{c}^{xy} + \mathbf{d} \\
& \bm{\beta} \ge 0 \\
\end{aligned} \end{equation}

\subsubsection{Formulation of a robust LP:}
Let $b_0 \in \mathbb{R}$ be a scalar value. We now define the following LP:

\begin{equation} \label{eq:lin_prog_rob} \begin{aligned}
\find \quad & \bm{\beta} \in \Rv{4e}, b_0 \in \Rv{} \\
\minimize  \quad & -b_0 \\
\st \quad &\mathbf{G} \bm{\beta} = \mathbf{D}^{xy} \mathbf{c}^{xy} + \mathbf{d} \\
& \bm{\beta} \ge b_0 \\
\end{aligned} \end{equation}

We observe that if $b_0$ is positive, LP~\ref{eq:lin_prog} admits a solution, and $b_0$ is proportional to the minimum distance of the equilibrium contact forces to the boundaries of the friction cones.
If $b_0$ is negative, the configuration is not in static equilibrium, and the value $b_0$ indicates "how far" from equilibrium
the configuration is. $b_0$ can thus be seen as a measure of robustness.

In our implementation, we solve the dual of LP~\ref{eq:lin_prog_rob}.
after a trivial elimination of the Lagrange multipliers associated to the inequality constraints, the dual problem has a lower dimensionality:
\begin{equation} \label{eq:dual} \begin{aligned}
\find \quad & \bm{\nu} \in \Rv{6}\\
\maximize  \quad & -(\mathbf{D}^{xy} \mathbf{c}^{xy} + \mathbf{d})^T \nu \\
\st \quad &\mathbf{G}^T \bm{\nu} \ge 0 \\
& \mathbf{1}^T \mathbf{G}^T \bm{\nu} = 1 \\
\end{aligned} \end{equation}

Indeed, from Slater's condition \citep{Boyd:2004:CO:993483}, we know that the optimal values of an LP and its dual are equal. Thus the optimal value of LP~\ref{eq:dual} is indeed the optimal $b_0$.

In conclusion, LP~\ref{eq:dual} provides a method to assert rapidly static equilibrium, and returns a robustness value.
This robustness criterion can be immediately transformed into a heuristic, and eventually be ponderated with others.
Another option to enforce robustness is to set a minimum acceptable value for $b_0$, but as a result this conservative approach is not complete.