### 2015-12-07 body/appendix_roms.tex
appendix

\section{Generating the $W$ volumes for HRP-2}
\label{app:rom}

In this Appendix,
we detail our generic method to generate the volumes $W$ used
in RB-RRT; with the example of HRP-2.