### 2015-12-07 main.tex

% sage_latex_guidelines.tex V1.01, 11 June 2015

\documentclass[shortAfour,sageh]{sagej}

\setcounter{secnumdepth}{3}
\usepackage{moreverb,url}
\usepackage{makeidx}         % allows index generation
\usepackage[percent]{overpic}
\usepackage{array} 
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{listings}
\usepackage{color}
\usepackage{bm}

\input{./math_commands.tex}

\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\def\volumeyear{2015}

\begin{document}

\runninghead{Tonneau et al.}

\title{A demonstration of the \LaTeXe\ class file for
\itshape{SAGE Publications}}

\author{Steve Tonneau\affilnum{1}, Andrea Del Prete\affilnum{1}, Julien Pettr\'e\affilnum{2}, Franck Multon\affilnum{3}, Nicolas Mansard\affilnum{1}}

\affiliation{\affilnum{1}LAAS-CNRS, Toulouse, France\\
\affilnum{2}Inria, Rennes, France \\
\affilnum{3}Universit\'e Rennes 2, France}

\corrauth{Steve Tonneau, LAAS-CNRS
7 av. Colonel Roche, \\
BP 54200,
31031 Toulouse cedex 4,
France.}

\email{pro@stevetonneau.fr}

\begin{abstract}
	\input{body/abstract}
\end{abstract}

\keywords{Class file, \LaTeXe, \textit{SAGE Publications}}

\maketitle

\section{Introduction}
\input{body/intro}
\input{body/overview}
\input{body/rbprm}
\input{body/contact}
\input{body/equilibrium}
\input{body/efort}
\input{body/results}
\input{body/conclusion}
\theendnotes
\bibliographystyle{SageH}
\bibliography{biblio}
\appendix
\input{body/appendix_algo}
\end{document}