### 2015-12-08 body/appendix_hpp.tex
appendix done, missing all stats and truck

\section{Source code of our planner}
\label{app:hpp}
Our planner is implemented using the Humanoid Path Planner (HPP) software.
HPP is an open source motion planning framework developed by the Gepetto team at LAAS-CNRS, that can be found at \url{http://projects.laas.fr/gepetto/index.php/Software/Main}.
The HPP modules implement the standard tools and algorithms used in motion planning,
such as the Bi-RRT planner from which RB-RRT is derived.

The robot models used in our experiments are described using the standard urdf file format, that is 
compatible with HPP.

Our implementation of the planner is also open source, and can be found at \url{https://github.com/stonneau/hpp-rbprm}. 
Contrary to the mature HPP, at the time of submitting this article,
our planner is not officialy released, mostly for lack of a complete documentation and a tutorial.

However a reader familiar with HPP should be able to use the library.
Our goal is to make our code useful for the community, and we are working on releasing a stable 
version of our planner as soon as possible.