### 2016-01-18 body/glossary.tex
nico + glossary

\newglossaryentry{quasi-flat}
{
  name=\textit{quasi-flat},
  description={refers to a contact surface for which the friction cone includes the direction opposite to the gravity}
}

\newglossaryentry{cluttered}
{
  name=cluttered,
  description={A class of contact planning problems that admit as a solution a sequence of contact configurations for which at least one contact occurs with a \gls{quasi-flat} surface}
}

\newglossaryentry{ccontact}
{
  name=$C_{Contact}$,
  description={Set of collision free full body configurations such that at least one end-effector is in contact with the environment}
}

\newglossaryentry{ccontact0}
{
  name=$C_{Contact}^0$,
  description={Set of \textit{contact feasible} root configurations $\mathbf{q}^{0}$, that verify $\exists \mathbf{q}^{\overline{0}},  \mathbf{q}^{0} \oplus \mathbf{q}^{\overline{0}} \in C_{contact}$}
}

\newglossaryentry{contact feasible}
{
  name=\textit{contact feasible},
  description={Refers to a root configuration that can be extended into a collision free contact configuration}
}

\newglossaryentry{equilibrium feasible}
{
  name=\textit{equilibrium feasible},
  description={Refers to a root configuration that can be extended into a collision free configuration in static equilibrium}
}

\newglossaryentry{ccequil}
{
  name=$C_{equil}$,
  description={Set of collision free full body configurations that are in static equilibrium}
}

\newglossaryentry{ccequil0}
{
  name=$C_{equil}^0$,
  description={Set of \textit{equilibrium feasible} root configurations $\mathbf{q}^{0}$, that verify $\exists \mathbf{q}^{\overline{0}},  \mathbf{q}^{0} \oplus \mathbf{q}^{\overline{0}} \in C_{equil}$}
}

\newglossaryentry{reachc}
{
  name=\textit{reachability condition},
  description={A root configuration $\mathbf{q}^{0}$ that verifies :\\ $W(\mathbf{q}^{0}) \cap O \neq \emptyset \wedge W^0_{s^*}(\mathbf{q}^{0}) \cap O = \emptyset$,
  where $s^*$ is a user-defined value}
}

\newglossaryentry{creach0}
{
  name=$C_{reach}^0$,
  description={Set of root configurations  $\mathbf{q}^{0}$ that verify the \gls{reachc}}
}

\newglossaryentry{interactive}
{
  name=interactive,
  description={We say that a planning method is \gls{interactive} when the computation time for one step is lesser than the
time to execute it. We arbitrarily approximate this time to one second.}
}