### 2017-01-11 body/overview.tex
Review Overview

% !TEX root =  ../main.tex
\section{Overview}
\label{overview}

Figure~\ref{fig:framework} illustrates our workflow.
\Pa and \Pb are addressed sequentially: we first plan a root guide path \deladp{with the \textit{reachability condition}}, before
extending it into a sequence of configurations in static equilibrium.

%
\begin{figure*}
  \centering
  \begin{overpic}[width=0.8\linewidth]{figures/contact_gen}
		\put (1,1) {a} 
		\put (22,1) {b} 
		\put (42,1) {c} 
		\put (62,1) {d} 
		\put (83,1) {e} 
	\end{overpic}
  \caption{Generation of a contact configuration for the right leg of HRP-2. (a): Selection of reachable obstacles. (b): Entries of the limb samples database (with $N = 4$). (c): With a proximity query on the octree database, configurations too far from obstacles are eliminated. (d): The best candidate according to a user-defined heuristic $h$ is chosen. (e): The final contact is achieved using inverse kinematics.}
  \label{fig:contact_gen}
\end{figure*}

\subsection{Computation of a guide path --- \Pa}
We first consider the problem of planning a root guide path (Figure~\ref{fig:framework}--$\mathcal{P}_1$).  The dimension of the path is equal to the number of degrees of freedom of the root of the robot.
\repadp{Similarly to}{As for} previous work~\citep{Bouyarmane2009} the path must be \textit{equilibrium feasible}: the root configurations \deladp{of the path} must be \repadp{extensible}{extended} into full-body static equilibrium configurations. Previous work verifies \textit{equilibrium feasibility} by explicitly computing such a \deladp{joint} configuration. 
We preserve the low dimensionality of the problem by approximating \textit{equilibrium feasibility} with \textit{contact rechability}, illustrated in the following.

An intuitive description of \gls{contact reachable} configurations is ``close, but not too close'' \addadp{to the environment}: close, because a contact surface must be partially included in the reachable workspace of the robot (represented for the right leg in Figure~\ref{fig:contact_gen}--a) to allow contact creation; not too close, because the robot must avoid collision.
\deladp{\textit{Contact reachability} is verified with the \textit{reachability condition}, a geometrical criterion based on this intuitive description (Figure~\ref{fig:reach_cond}).}
More precisely, a root configuration is \textit{contact reachable} if the root scaled by a \addadp{user-defined} factor $s \geq 1$ is not in collision, while the reachable workspace is in collision.
%~ Verifying the \textit{reachability condition} is very efficient in practice.
To plan a root path, we then use an RRT planner where, instead of checking collision to validate a configuration, we verify the \textit{reachability condition}.
This is detailed in Section~\ref{rbprm}.
%~ In Section~\ref{rbprm} we detail how  we define and use the \textit{reachability condition} to compute a guide path with the Reachability-Based RRT (Figure~\ref{fig:framework}---\Pa).

\begin{figure}[t]
\centering
  \begin{overpic}[width=0.8\linewidth]{figures/reachability_cond}
	\end{overpic}
\caption{The reachability condition is verified by the right configuration: the trunk (red) is free of collisions while the limbs reachable workspace (green) intersect the environment.}
		   \label{fig:reach_cond}
\end{figure}

%
%~ Ideally we would like to sample root configurations that are \gls{equilibrium feasible}. A root configuration is \gls{equilibrium feasible} if and only if there exists at least one joint configuration such that:
%~ \begin{description}
%~ \item[C1]: the robot is not in collision with the environment, 
%~ \item[C2]: the robot is in static equilibrium. 
%~ \end{description}
%~ Verifying C2 is computationally expensive because it requires the projection of the whole-body configuration into a contact posture, as well as the resolution of a linear program to compute the contact forces~\citep{Prete2016}.
%~ For this reason, we are interested in finding an approximation of the \glslink{equilibrium feasible}{\textit{equilibrium feasibility}} that is computationally efficient.
%~ A requirement for C2 is the condition:
%~ \begin{description}
%~ \item[C3]: at least one end-effector is in contact with the environment.
%~ \end{description}
%~ C1 and C3 define \glslink{contact reachable}{\textit{contact reachability}}, which is easier to verify than \glslink{equilibrium feasible}{\textit{equilibrium feasibility}} because it does not require the computation of contact forces.
%~ We make the assumption that for the \textit{\gls{cluttered} contact planning problem}, most of the times \gls{contact reachable} root configurations are also \gls{equilibrium feasible}.
%~ For this reason we neglect C2 during the resolution of $\mathcal{P}_1$, and we consider it only when solving \Pb.
%~ However, even verifying \glslink{contact reachable}{\textit{contact reachability}} is computationally too expensive because it requires the projection of the whole-body configuration into a contact posture.
%~ 
%~ An intuitive description of \gls{contact reachable} configurations is ``close, but not too close'': close, because a contact surface must be partially included in the reachable workspace of the robot (represented for the right leg in Figure~\ref{fig:contact_gen}--a); not too close, because the robot must avoid collision.
%~ We approximate \glslink{contact reachable}{\textit{contact reachability}} with the \textit{reachability condition}, a geometrical criterion based on this intuitive description.
%~ More precisely, a root configuration is \textit{reachable} if the root scaled by a factor $s \geq 1$ is not in collision, while the reachable workspace is in collision with the environment.
%~ Verifying the \textit{reachability condition} is very efficient in practice.
%~ In Section~\ref{rbprm} we detail how  we define and use the \textit{reachability condition} to compute a guide path with the Reachability-Based RRT (Figure~\ref{fig:framework}---\Pa).

In the remainder of this paper, we use the terms \gls{contact reachable} and \gls{equilibrium feasible} to qualify either a root configuration, a whole-body configuration, or a set
of such configurations.

\subsection{Generating a discrete sequence of contact configurations --- \Pb}

The second stage extends the guide path into a sequence of contact configurations (Figure~\ref{fig:framework}--$\mathcal{P}_2$).
To achieve this the root path is first decomposed into a sequence of discrete root configurations, according to a user-defined discretization step.
Each root configuration is then extended into a whole-body configuration, subject to three constraints:
\begin{itemize} 
\item At most one contact is not maintained (\textit{broken}) between two consecutive configurations.
\item Each configuration is in static equilibrium.
\item Each configuration is collision-free.
\end{itemize}

The algorithm thus proceeds iteratively, starting from the whole-body initial configuration of the robot.

The algorithm takes advantage of the fact that each root configuration is fixed to generate the contact by considering each limb individually. Candidate limb configurations resulting in the end-effector being close to the environment are selected from a database (\repadp{computed offline independently}{independent} from the environment).
Then we select one of these configurations based on user-defined heuristics (Figure~\ref{fig:contact_gen}), detailed in Appendix~\ref{sec:heuristics}.

The details of the contact sequence generation are given in Section~\ref{sec:contact}.

%~ From a given start configuration, the planner proceeds in an iterative fashion along the discretized path: given a new root configuration, an inverse-kinematics solver 
%~ is used to maintain the contacts that were previously existing. Possibly, some of these contacts cannot be maintained (because of joint limits or collisions).
%~ Then the contacts are broken. Conversely, new contacts are created to ensure the static equilibrium of the robot.
%~ The algorithm is designed so that only one contact can be created or broken between two successive configurations. While it does not provide guarantees that the interpolation between two configurations is achievable, it appears empirically as a reasonable heuristic.

%~ To create a contact, we consider each limb of the robot as a manipulator arm attached to the root. We store a database of configurations for each manipulator. Figure~\ref{fig:contact_gen} -- 2 presents a few configurations for the right arm of the robot. The configurations in the database which are close to contact are considered (Figure~\ref{fig:contact_gen} -- 2 and 3). Among the candidates, 3 criteria are considered to choose the most relevant: the ideal candidate is free of collision, allows to maintain static balance if necessary, and  contributes to the motion in an optimal way. We call the measure of this contribution task efficiency: it represents the ability the limb has to exert a force compatible with the direction of motion. In Figure~\ref{fig:contact_gen} -- 4, the black arrow represents the direction of motion. Among the candidates, the most relevant to achieve a vertical motion is chosen, and projected on the contact surface using an inverse kinematics solver (Figure~\ref{fig:contact_gen} -- 5). The task efficiency is measured using the Extended FORce Transmission ratio (EFORT), proposed by ~\cite{Tonneau:2014:TEC:2619648.2619652}.

\begin{figure}
  \centering
  \includegraphics[width=0.95\linewidth]{figures/HyQ_roms}
  \caption{
           Reachable workspace and torso bounding box of HyQ. The green shapes represent the reachable workspace $W^k$ of each limb. The red shape is $W^0$.}
		   \label{fig:HyQ_roms}
\end{figure}

\section{Notation \deladp{conventions} and definitions} \label{notations}

A vector  $\mathbf{x}$ is denoted with a bold lower-case letter.
A matrix $\mathbf{A}$ is denoted with a bold upper-case letter.
A set $C$ is denoted with an upper-case italic letter.
Scalar variables and functions are denoted with lower-case italic letters, such as
$r$ or $f(\textbf{x})$.

%\subsubsection{Robot description}
%\label{def}

%~ \begin{figure}
  %~ \centering
  %~ \begin{overpic}[width=0.8\linewidth]{figures/character}
    %~ \put (7,29) {\small{$\in \mathbf{q}^1$}}
    %~ \put (7,25) {\small{effector}}
  %~ \end{overpic}
  %~ \caption{
    %~ Left: Robot in a rest configuration. The right arm is denoted as the limb $R^1$. Each colored dot represents a degree of freedom around an axis. Right: Volumes of the robot. The red geometry denotes $W^0$ and must remain collision-free. The green spheres are the reachable workspace of each limb, the  $W^k$.}
  %~ \label{fig:character}
%~ \end{figure}

\medskip
\textbf{A robot} is a kinematic chain $R$ composed of: a root $R^0$, and $l$ limbs $R^k, 1 \leq k \leq l$, attached to the root.
The root has $r \geq 6$ DOFS: for instance, HRP-2 has two extra DOFS in the torso, such that we have $r=8$.
Therefore $R$ is fully described by a configuration $\mathbf{q} \in  \mathbb{R}^{r+n}$.
$\mathbf{q}$ is decomposed as follows:
\begin{itemize}
	\item $\mathbf{q}^k$ is a configuration (a vector of joint values) of the limb $R^k$; % (Figure~\ref{fig:character});
	\item $\mathbf{q}^{\overline{k}}$ is a vector of joint values of R \textbf{not} related to $R^k$. We define for convenience \mbox{$\mathbf{q}= \mathbf{q}^k \oplus \mathbf{q}^{\overline{k}}$}; %Abusively, we denote by $\mathbf{q}^0 \oplus \mathbf{q}^k$ any configuration respecting $\mathbf{q}^0$ and $\mathbf{q}^k$;
	\item $\mathbf{q}^{0}\in \mathbb{R}^r$ is the world coordinates vector of $R^0$.
\end{itemize}
%Finally, $ \mathbf{q}_{start}$ and  $\mathbf{q}_{goal}$ are the start and goal configurations given as an input of our problem.

%\subsubsection{Environment and other volumes} 
%\label{sec:A}
\medskip
\noindent $W^0$ is a volume encompassing $R^0$ (Figure~\ref{fig:HyQ_roms}). \\
 %~ (Figure~\ref{fig:character}-right: central cylinder). 
 $W^k$ is the reachable workspace of a limb $R^k$: %(Figure~\ref{fig:character}-right: the four ellipses). 
\begin{equation}
  W^k = \left\{ {\mathbf{x} \in \mathbb{R}^3: \exists \, \mathbf{q}^k \in C^k_{jl}, \mathbf{p}^k(\mathbf{q}^k) = \mathbf{x} } \right\}
\end{equation}
%~ \adnote{in this definition you don't account for joint limits}
where $\mathbf{p}^k$ denotes the end-effector position of $R^k$ (translation only) for $\mathbf{q}^0 = \mathbf{0}$ being the null displacement, and  $C^k_{jl}$ is the space
of admissible limb joint configurations. We also define \mbox{$W = \bigcup_{k=1}^{l}W^k$}, and
$W^k(\mathbf{q}^{0})$ (for $1 \leq k \leq l$) as the volume $W^k$ for the root configuration $\mathbf{q}^{0}$.

\medskip
\textbf{The environment} $O$ is defined as the union of the obstacles $O_i$ that it contains.

\medskip
$C_{Contact}$ is the set of whole body-configurations in contact and collision-free.

\medskip
$C_{Equil} \subset C_{Contact}$ is the set of whole body-configurations in static equilibrium and collision-free.

\medskip
For any set $C_{X}$, we define $C_{X}^0$:
\begin{equation*}
  C_{X}^0 = \left\{ \mathbf{q}^{0},  \exists \mathbf{q}^{\overline{0}}: \mathbf{q}^0  \oplus \mathbf{q}^{\overline{0}} \in C_{X} \right\}
\end{equation*}

%~ Finally, we define $dist(\mathbf{x}, O)$ as the minimal distance between a point $\mathbf{x}$ and the closest surface of the environment.