### 2017-01-20 body/appendix_optim.tex
updated appendix optim

\section{Addressing $\mathcal{P}_3$}
\label{app:optim}
The stair climbing and the standing up scenarios were validated with the trajectory optimization scheme provided in \citeauthor{Carpentier2016}. 
The optimization is formulated as an optimal control problem that aims at generating a valid trajectory for the center of mass.
To address the others scenarios, we propose a new implementation, entirely open source (\url{https://github.com/stonneau/python_sandbox}), which can be integrated directly in our motion planner software. This formulation uses the center of mass acceleration and angular momentum as input variables, while previously the cotnact forces were used.
The center of mass trajectory resulting from the optimization is then turned into a collision-free whole-body trajectory.

We rewrite Eq.~\ref{eq:new_eul} in the general case:

%~ \begin{align} \label{eq:new_eul}
%~ \underbrace{
%~ \mat{\mathbf{I}_3 & \dots & \mathbf{I}_3 \\
%~ \hat{\mathbf{p}}_1 & \dots & \hat{\mathbf{p}}_e} \mathbf{V}
%~ }_\mathbf{G} \bm{\beta}, = 
%~ \underbrace{\mat{\mathbf{0}_{3\times 3} \\ m \hat{\mathbf{g}}}}_{\mathbf{D}} \mathbf{c} + 
%~ \underbrace{\mat{-m\mathbf{g} \\ \mathbf{0}}}_{\mathbf{d}}
%~ \end{align}

\begin{align} \label{eq:new_eul_acc}
\mathbf{G} \bm{\beta} = 
\underbrace{\mat{m (\mathbf{\ddot{c}} - \mathbf{g}) \\m (\mathbf{\ddot{c}} - \mathbf{g}) + \dot{L}}}_{\mathbf{w}}
%~ \underbrace{m (\mathbf{\ddot{c}} - \mathbf{g})  \\ m (\mathbf{\ddot{dc}} - \mathbf{g}) + \mathbf{\dot{L}} }_{\mathbf{w}}
\end{align}

where $\dot{L}$ is the angular momentum expressed at the com.
Eq.~\ref{eq:new_eul_acc} defines a 6 dimensional cone $\mathcal{K}$ ~\citep{qiu:dhm:2011,Caron2015}. For a given set of contacts,
this cone determines all the contact forces that generate an admissible wrench $\mathbf{w}$ such that the robot won't fall.
The face form of $\mathcal{K}$ can be computed using the double description method~\citep{Fukuda1996}, resulting in the following linear inequalities:

\begin{equation}
\label{eq:cone_k}
	\mathcal{K} =  \left\{ \mathbf{w}, \mathbf{A}\mathbf{w} \leq \mathbf{b	} \right\}
\end{equation}

The objective 
is then to plan a trajectory for the center of mass such that the generated $\mathbf{w}$ always verifies Eq.~\ref{eq:cone_k}. 
We now consider two contact configurations  $\mathbf{q}_0$ and $\mathbf{q}_1$ computed by our planner: in the general case one contact is broken and one created to get from
$\mathbf{q}_0$ to $\mathbf{q}_1$. The motion is thus constrained by three cones $\mathcal{K}_s, s =0 \dots 2$. We manually define the duration of each of the three contact phases.
We call the total duration of the motion $T$, and formulate the following optimization problem:

\begin{equation} \label{eq:lin_prog_rob} \begin{aligned}
%~ \find \quad & \mx{R} \in \Rv{3\times T}, \mxdd{C} \in  \Rv{6\times T} \\
\minimize_{\mathbf{c}(t), \mathbf{w}(t)}  \quad & \sum_{s=0}^{2} \int_{t_s}^{t_s + \Delta t_s}  h(\mathbf{c}, \mathbf{w})  \\
\st \quad & \mathbf{\dot{c}} = f(\mathbf{c}, \mathbf{w}) \\
	\quad & \mathbf{A}_s \mathbf{w}(t)  \leq \mathbf{b}_s , \forall t \in  \left[t_s, t_s + \Delta t_s \right[ \\
	\quad & \mathbf{D}_s \mathbf{c}(t)  \leq \mathbf{d}_s , \forall t \in  \left[t_s, t_s + \Delta t_s \right[ \\
	\quad & \mathbf{c}(0)  = \mathbf{c}_{\mathbf{q}_0} \\
	\quad & \mathbf{c}(T)  = \mathbf{c}_{\mathbf{q}_1} \\
	\quad & \mathbf{c}(0)  = \mathbf{\dot{c}}(0)  = \mathbf{\ddot{c}}(0) = \mathbf{0}\\
	\quad & \mathbf{c}(T)  = \mathbf{\dot{c}}(T)  = \mathbf{\ddot{c}}(T) = \mathbf{0}\\
    %~ \quad & ({\vct{r}}\tra\vct{r}=1,  \forall t  ?) \\
    %~ \quad & \delta t \sum\limits_1^{T-1}{\vc{r}^1}\tra {\vc{r}^1} \ge t_{transit} \\
    %~ \quad & \mx{A}_{\int} \mxdd{C} \leq \vc{a}_{\int} \\
    %~ \quad & \mx{V}_{\int} \mxd{C} \leq \vc{v}_{\int} \\
    %~ \quad & \mx{B}_{\int}^{1} \mx{C} \leq \vc{b}_{\int} \\
    %~ \quad & [\vc{c}_0\tra, \vcd{c}_0\tra]\tra = \vc{x}_0 \\
    %~ \quad & \vc{c}_T = \vc{c}_T^{input} \\
\end{aligned} \end{equation}

As a cost function, $h$ typically minimizes the angular momentum and acceleration variation over the whole trajectory. $f$ is the integration method
that ties together the position, velocity and acceleration of the center of mass. 
$\mathbf{D}_s$ and $\mathbf{d}_s$ denote stacked kinematic constraints on the center of mass position, determined by the active contact locations.
The inequalities for each contact are determined in the same way that the reachable workspace is computed in Appendix~\ref{app:rom}, with
the effector serving as root (a triangulated object is a polytope, and as such can be described as a set of linear inequalities).

This formulation can trivially be extended over the whole contact sequence.
In our implementation, the problem is discretized using time steps of 100 ms.

As an output, the solution of this optimization problem gives us a admissible center of mass trajectory.
To compute the whole body motion, we use a two step approach.

First, we plan a kinematic motion for the robot, subject to the contact constraints. We also constrain the center of mass to follow
the computed trajectory. This is achieved using a constraint-based RRT planner~\citep{7759083}. As a result we obtain a whole-body motion that is collision free.

The entire resolution takes approximatively 1.5 seconds for a complete contact transition.

This motion is then validated in our dynamic simulator~\citep{DelPrete2015b}, using the trajectory as a reference postural task to an inverse dynamics
solver, that accounts for the torque constraints of the robot. This final result is shown in the videos.