### 2017-01-23 body/abstract.tex
andrea s last modif plus 250 words abstract

% !TEX root =  ../main.tex
We present a contact planning framework for complex legged locomotion tasks: standing up, climbing stairs using a handrail, crossing rubble and getting out of a car. The need for such a planner was highlighted at the Darpa Robotics Challenge, where multi-contact behaviors
could not be demonstrated.

Indeed, current planners suffer from prohibitive computation times, due to the difficulty
of generating contact postures: the feasible space (the contact manifold) is of high dimension,
and impossible to sample.

We tackle this issue by introducing the reachability condition. This condition allows us to define
a geometric approximation of the contact manifold, which is of low dimension and can be sampled efficiently.
Informally, the condition verifies that the root configuration of a robot is close, but not too close to obstacles: close to allow contact creation, not too close to avoid collision. Thanks to this condition we decompose the hard contact planning problem into two
sub-problems: first, planning a guide path for the root without considering the whole-body configuration, using a sampling-based algorithm; then, generating a discrete sequence of whole-body configurations in static equilibrium along this path, using a deterministic contact-selection algorithm.

Our approach results from the pragmatic choice of favoring efficiency over completeness, which we justify empirically: in a
few seconds, with high success rates, we generate complex contact plans for various scenarios and robots: HRP-2, HyQ, and a dexterous hand. These plans are then validated either in dynamic simulations, or on the real HRP-2 robot, thus demonstrating 
the first interactive multi-contact planner.